---
title: Interacting with the SewerRat REST API
author:
- name: Aaron Lun
  email: infinite.monkeys.with.keyboards@gmail.com
package: SewerRat
date: "Revised: April 24, 2024"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Hitting the SewerRat API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(BiocStyle)
self <- Githubpkg("ArtifactDB/SewerRat")
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
```

# Introduction

The `r self` package implements an R client for the [API of the same name](https://github.com/ArtifactDB/SewerRat).
This allows users to easily register or deregister their own directories in the index for subsequent queries.
Readers are referred to [SewerRat documentation](https://github.com/ArtifactDB/SewerRat) for a description of the concepts;
this guide will strictly focus on the usage of the `r self` package.
For demonstration purposes, we'll set up a test instance of the API on our local machine:

```{r}
library(SewerRat)
info <- startSewerRat()
info$url
```

# Registering directories

We assume that the users of the `r self` package and the SewerRat API itself are both on the same shared filesystem.
The idea is that a user can instruct SewerRat to incorporate a particular directory into the search index,
provided that the to-be-registered directory is world-readable and that the caller has write access to it.
To illustrate, let's mock up a directory of metadata files:

```{r}
mydir <- tempfile()
dir.create(mydir)
write(file=file.path(mydir, "metadata.json"), '{ "name": "foo", "description": "bar" }')
dir.create(file.path(mydir, "stuff"))
write(file=file.path(mydir, "stuff", "metadata.json"), '{ "food": "barramundi" }')
```

We can then easily register it with the `register()` function.
The example below will only index metadata files inside `mydir` named `metadata.json`, though any number of base names can be supplied here.

```{r}
library(SewerRat)
register(mydir, names="metadata.json", url=info$url) 
```

Similarly, we can deregister this directory with `deregister()`.

# Searching the index

We use the `query()` function to perform free-text searches on the indexed metadata.
This function does not require filesystem access and can be executed remotely.

```{r}
lapply(query("foo", url=info$url), function(x) x$path)
lapply(query("bar%", url=info$url), function(x) x$path) # partial match to 'bar...'
lapply(query("bar% AND foo", url=info$url), function(x) x$path) # boolean operations
lapply(query("food:bar%", url=info$url), function(x) x$path) # match in the 'food' field
```

We can also search on the user, path components, and time of creation:

```{r}
lapply(query(user=Sys.info()["user"], url=info$url), function(x) x$path) # created by myself
lapply(query(path="stuff", url=info$url), function(x) x$path) # path has 'stuff' in it
lapply(query(from=Sys.time() - 3600, url=info$url), function(x) x$path) # created less than 1 hour ago
```

Or indeed, any combination of these fields, which are treated as a boolean AND in the query:

```{r}
query("barramundi", path="stuff", url=info$url)
```

# Session information {-}

```{r}
sessionInfo()
```
