---
title: Interacting with the SewerRat REST API
author:
- name: Aaron Lun
  email: infinite.monkeys.with.keyboards@gmail.com
package: gypsum
date: "Revised: April 2, 2024"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Hitting the gypsum API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(BiocStyle)
self <- Githubpkg("ArtifactDB/SewerRat")
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
```

# Introduction

The `r self` package implements an R client for the [API of the same name](https://github.com/ArtifactDB/SewerRat).
This allows users to easily query the SewerRat search index, register or deregister their own directories in the index, and quickly save and load Bioconductor objects for registration.
It is assumed that the users of the `r self` client and the SewerRat API itself are accessing the same shared filesystem;
this is typically the case for high-performance computing (HPC) clusters in scientific institutions.

# Registering directories

# Searching the index

# Saving Bioconduct objects

# Session information {-}

```{r}
sessionInfo()
```
